---
title: "theft"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
```{r}
library(arm)
library(lme4)
library(ggplot2)
library(lmtest)
library(spatialreg)
library(RColorBrewer)
library(sjstats)
library(tseries)
```

## visualising monthly data as heat map.
```{r}
wmatrix <- read_csv("wmatrix.csv")
ward<-data.matrix(wmatrix)
coul1<-colorRampPalette(brewer.pal(9,"BuPu"))(25)
heatmap(ward,Rowv=NA,Colv=NA,scale="none",col=coul1)
```

## KPSS test for monthly data. Considering which data frequency to use for modeling, monthly or annual.
```{r}
s1<-read.csv("series1.csv",header=F)$V1
kpss.test(s1)
pp.test(s1)
```


## fitting model with 2018 theft volume and time lagged predictors and examine model fit, calculate ICC.
```{r}
test18 <- read_csv("test518c.csv")
m0<-lmer(Theft18~1+(1|boro),data=test18)
m1<-lmer(Theft18~1+(1|boro)+(1|cl),data=test18)
summary(m1)
coef(m1)$boro
fixef(m1)
ranef(m1)$boro
se.fixef(m1)
se.ranef(m1)$boro
coef(m1)$boro[32,1]+c(-2,2)*se.ranef(m1)$boro


m2<-lmer(Theft18~1+inc+oldpop+trans+(1|boro),data=test18)
summary(m2)
plot(m2)
u=tapply(test18$trans,test18$boro,mean)
a.hat.m2=fixef(m2)[1]+fixef(m2)[4]*u+ranef(m2)$boro[,1]
a.se.m2=se.ranef(m2)$boro[,1]
lower=a.hat.m2-a.se.m2
upper=a.hat.m2+a.se.m2
par(mar=c(3.5,3.5,1,1),mgp=c(2.5,.7,0),cex.lab=1.5,tck=-.01,cex=1.2,las=1,cex.axis=1.5,pch=20)
par(mar=c(3.5,3.5,1,1),mgp=c(2.5,.7,0),cex.lab=1.5,tck=-.01,cex=1.2,las=1,cex.axis=1.5,pch=20)
plot(u,a.hat.m2,ylim=range(lower,upper),xlab="borough",ylab="est.intercept")
curve(fixef(m2)[1]+fixef(m2)[4]*x,lwd=1.5,col="black",add=TRUE)


icc(m2, adjusted = FALSE)
```

## model comparison with key predictors. 
## Results suggest random intercept model, as there's not sufficient information to estimate random slopes.
```{r}
test18 <- read_csv("test518c.csv")
m2<-lmer(Theft18~1+inc+oldpop+trans+(1|boro),data=test18)
m3<-lmer(Theft18~1+inc+oldpop+rn+(1|boro),data=test18)
m4<-lmer(Theft18~1+inc+oldpop+pop+rn+(1|boro),data=test18)
m5<-lmer(Theft18~1+mp+oldpop+pop+rn+(1|boro),data=test18)
m6<-lmer(Theft18~1+mp+oldpop+pop+rn+mp:rn+(1+mp|boro),data=test18)
m7<-lmer(Theft18~1+mp+oldpop+pop+(1|boro)+(1|cl),data=test18)
summary(m2)
summary(m3)
summary(m4)
summary(m5)
summary(m6)
summary(m7)
plot(m2)
plot(m3)
plot(m4)
plot(m5)
plot(m6)
plot(m7)

```


## with chosen formula, test model prediction of 2019 with real 2019.

```{r}
test19 <- read_csv("test19.csv")
str(p2019<-predict(m7,test19))
p2020
```


## with chosen formula, fit model with 2017-2019data for 2020 prediction
```{r}
fit19 <- read_csv("fit19.csv")
test20 <- read_csv("test20c.csv")
m72<-lmer(Theft18~1+mp+oldpop+pop+(1|boro)+(1|cl),data=fit519)
display(m72)
plot(m62)
str(p2020<-predict(m72,test20))
p2020
```

